This is the title [Article v\versionnumber]

### 2022-08-19 manuscript/manuscript.tex

\subsection{Software/system requirements}
%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.). Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
The main implementation of GIST is part of the MD analysis software \software{cpptraj} as part of AmberTools \cite{amber22}.
A recent version of \software{cpptraj} should be used, such as the one in AmberTools22 or an installation from GitHub.

The closed-source Amber simulation engine is used in this tutorial to run the necessary MD simulations.
If no Amber installation is available, the MD can be performed other simulation engines compatible with \software{cpptraj}.
If using GROMACS, we propose to prepare the structures and topologies with AmberTools and converting them using \software{acpype.py}.
In that way, a full Amber topology is available for GIST.
Also note that compressed trajectory formats such as \inlinecode{xtc} might bias the entropy calculation.

Furthermore, a recent Python version (>3.6) should be available with the following packages.
\begin{itemize}
	\item \software{gisttools} (and its prerequisites)
	\item \software{mdtraj}
\end{itemize}

@@

If the reader prefers to skip the MD and GIST calculation, we also provide the GIST output files with this tutorial, such that the post-processing can be done without any expensive calculations. We also provide a Jupyter notebook \cite{Kluyver2016-jupyter,Granger2021-jupyter} with the presented analyses.

\section{Theory}
GIST is an implementation of inhomogeneous fluid solvation theory (IST) \cite{Lazaridis1998} that discretizes the free energy of solvation $\dgsolv$ on a three-dimensional grid. 
It was first devised by Nguyen et al. \cite{Nguyen2012} and its implementation in \software{cpptraj} was thoroughly described by Ramsey et al. \cite{Ramsey2016}. % to overcome the limitation of IST which was commonly used for regions of high water occupancy and its implementation in \software{cpptraj} was thoroughly described in \cite{Ramsey2016}.
Here, we only provide a short overview of the theory behind GIST.
For more detailed information, we recommend one of the more recent publications on developments in GIST. \cite{Kraml2020,Chen2021}

\subsection{Solvation Entropy}
%The solvation entropy of a system is totality of the contribution of the  solute-water entropy and the water-water entropy. -- this is inconsistent with IST theory (fwa)

%\begin{equation}
%	\Delta S_\textit{solv} = \Delta S_\textit{sw} + \Delta S_\textit{ww}
%\end{equation}
%Here the solvation entropy is approximated from the contribution of solute-water interaction. 
%\textbf{STILLNEEDS REVISON + ADD CITATIONS} The quantity $k_\textit{b}$represents the Boltzman constant, 
IST expresses the free energy of solvation in terms of the solvent density in a coordinate system defined by the solute.
The solute is kept fixed in space, following the standard solvation process established by Ben-Naim \cite{ben-naim-book}.

@@

The solvent density is expressed relative to the bulk density $\rho^\textit{o}$. %$g_\textit{sw}\left(\textbf{r},\omega \right)$ is the solute-water 
%pair correlation function in the solute's frame of reference, \textbf{r} is the coordinates of the oxygen atom of the water molecule,  
%$\omega$ represents the Euler angles, and 
The entropy integral is also normalized by $8\pi^2$, which is the volume of the orientational space.
In bulk or at a high distance to the solute, the quantity $g_\textit{sw}\left(\textbf{r},\omega \right)$ is unity which leads to the first-order entropy of 
bulk to be zero. 
Therefore, no reference entropy needs to be computed unless there is a numeric bias, e.g., when the MD frames are not statistically independent due to insufficient spacing.
%Also, this quantity reaches unity as the distance from the solute molecule increases which leads for the calculation of 
%the solvation entropy to be an approximation of the integral around the solute atom. 
\begin{equation}
	\Delta S_\textit{solv}
	\approx \Delta S_\textit{sw}
	\newcommand{\gro}{g_{sw}(\textbf{r},\omega)}
	\equiv -k_\textit{B} \frac{\rho^\textit{0}}{8\pi^\textit{2}} \int \gro \ln\gro d\textbf{r}d\omega
\end{equation}
\todo{We are inconsistent regarding R/kB}

@@

The average entropy contributions per solvent molecule in voxel $k$ are calculated as
\begin{equation}
	S_{k}^\textit{trans} = -R \left( \gamma + \frac{1}{N_\textit{k}} \sum _{i=1}^{N_k} \ln g_{NN, \textit{i}}(\textbf{r}) \right)
\end{equation}
and
\begin{equation}
	S_{k}^\textit{orient} = -R \left( \gamma + \frac{1}{N_k} \sum _{i=1}^{N_k} \ln g_{NN, i}(\omega) \right)
	,
\end{equation}
where $N_\textit{k}$ is the number of water molecules found in voxel k, $\gamma$ is Euler's constant that accounts for the bias in naive entropy estimator, and $g_\textit{NN}$ is the nearest neighbor estimate.

@@

\begin{equation}
	S_\textit{k} = -R \left( \gamma + \frac{1}{N_\textit{k}} \sum _{i=1}^{N_k} \ln g_{NN, \textit{i}}(\textbf{r},\omega) \right)
\end{equation}

>>>>

\subsection{Software/system requirements}
%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.). Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
The main implementation of GIST is part of the MD analysis software \software{cpptraj} as part of AmberTools \cite{amber22}.

A recent version of \software{cpptraj} should be used, such as the one in AmberTools22 or an installation from GitHub (\url{https://github.com/Amber-MD/cpptraj}).
The closed-source Amber simulation engine is used in this tutorial to run the necessary MD simulations.

The closed-source Amber simulation engine is used in this tutorial to run the necessary MD simulations.
If no Amber installation is available, the MD can be performed other simulation engines compatible with \software{cpptraj}.
If using GROMACS, we propose to prepare the structures and topologies with AmberTools and converting them using \software{acpype.py}.
In that way, a full Amber topology is available for GIST. 
\todo{ (vah) is it necessary to mention GROMACS specifically? I would remove the previous two lines}
Also note that compressed trajectory formats such as \inlinecode{xtc} might bias the entropy calculation.

Furthermore, a recent Python version (>3.6) should be available with the following packages.
\begin{itemize}
	\item \software{gisttools} (\url{https://github.com/liedllab/gisttools})
	\item \software{mdtraj (\url{https://www.mdtraj.org/})}
\end{itemize}

@@

If the reader prefers to skip the MD and GIST calculation, we also provide the GIST output files with this tutorial, such that the post-processing can be done without any expensive calculations.
We also provide a Jupyter notebook \cite{Kluyver2016-jupyter,Granger2021-jupyter} with the presented analyses. 
We recommend JupyterLab to work with the notebook. Jupyter is available at \url{https://jupyter.org/}.

For the visualization of three-dimensional contributions we require molecular visualization software that can handle \software{OpenDX} files.
We recommend using a current version of \software{PyMol}\cite{pymol} or \software{VMD}\cite{vmd}.
\section{Theory}

GIST is an implementation of inhomogeneous fluid solvation theory (IST) \cite{Lazaridis1998} that discretizes the free energy of solvation $\dgsolv$ on a three-dimensional grid. 
It was first devised by Nguyen et al. \cite{Nguyen2012} and its implementation in \software{cpptraj} was thoroughly described by Ramsey et al. \cite{Ramsey2016}. % to overcome the limitation of IST which was commonly used for regions of high water occupancy and its implementation in \software{cpptraj} was thoroughly described in \cite{Ramsey2016}.
Here, we only provide a short overview of the theory behind GIST.
For more detailed information, we recommend one of the more recent publications on developments in GIST. \cite{Kraml2020,Chen2021}

\subsection{Solvation Entropy}

%The solvation entropy of a system is totality of the contribution of the  solute-water entropy and the water-water entropy. -- this is inconsistent with IST theory (fwa)

%\begin{equation}
%	\Delta S_\textit{solv} = \Delta S_\textit{sw} + \Delta S_\textit{ww}
%\end{equation}
%Here the solvation entropy is approximated from the contribution of solute-water interaction. 
%\textbf{STILLNEEDS REVISON + ADD CITATIONS} The quantity $k_\textit{b}$represents the Boltzman constant, 
%$\rho^\textit{o}$ is the bulk number density, $g_\textit{sw}\left(\textbf{r},\omega \right)$ is the solute-water 
%pair correlation function in the solute's frame of reference, \textbf{r} is the coordinates of the oxygen atom of the water molecule,  
%$\omega$ represents the Euler angles, and $\frac{1}{8\pi^2}$ is the volume of the orientational space, which acts as a normalization constant.
%In bulk and in regions where the orientational 
%entropy is uniform, the quantity $g_\textit{sw}\left(\textbf{r},\omega \right)$ is unity which leads to the first-order solvation entropy of 
%bulk to be zero. Also, this quantity reaches unity as the distance from the solute molecule increases which leads for the calculation of 
%the solvation entropy to be an approximation of the integral around the solute atom.

IST expresses the free energy of solvation in terms of the solvent density in a coordinate system defined by the solute.
The solute is kept fixed in space, following the standard solvation process established by Ben-Naim \cite{ben-naim-book}.

@@

The solvent density is expressed relative to the bulk density $\rho^\textit{o}$. %$g_\textit{sw}\left(\textbf{r},\omega \right)$ is the solute-water 
%pair correlation function in the solute's frame of reference, \textbf{r} is the coordinates of the oxygen atom of the water molecule,  
%$\omega$ represents the Euler angles, and 
The entropy integral is also normalized by $8\pi^2$, which is the volume of the orientational space.
In bulk or at a high distance to the solute, the quantity $g_\textit{sw}\left(\textbf{r},\omega \right)$ is unity which leads to the first-order entropy of 
bulk to be zero. 
Therefore, no reference entropy needs to be computed unless there is a numeric bias, e.g., when the MD frames are not statistically independent due to insufficient spacing.
%Also, this quantity reaches unity as the distance from the solute molecule increases which leads for the calculation of 
%the solvation entropy to be an approximation of the integral around the solute atom. 
\begin{equation}
	\Delta S_\textit{solv}
	\approx \Delta S_\textit{sw}
	\newcommand{\gro}{g_{sw}(\textbf{r},\omega)}
	\equiv -R \frac{\rho^\textit{0}}{8\pi^\textit{2}} \int \gro \ln\gro d\textbf{r}d\omega
\end{equation}
\todo{We are inconsistent regarding R/kB; vah: Fixed. is it - tho? Papers seem to disagre }

@@

The average entropy contributions per solvent molecule in voxel $k$ are calculated as

\begin{equation}
	S_{k}^\textit{trans} = -R \left( \gamma + \frac{1}{N_\textit{k}} \sum _{i=1}^{N_k} \ln g_{NN, \textit{i}}(\textbf{r}) \right)
\end{equation}
and
\begin{equation}

	S_{k}^\textit{orient} = -R \left( \gamma + \frac{1}{N_k} \sum _{i=1}^{N_k} \ln g_{NN, i}(\omega) \right)
	,

\end{equation}
where $N_\textit{k}$ is the number of water molecules found in voxel k, $\gamma$ is Euler's constant that accounts for the bias in naive entropy estimator, and $g_\textit{NN}$ is the nearest neighbor estimate.

@@

\begin{equation}

	S_\textit{k} = -R \left( \gamma + \frac{1}{N_\textit{k}} \sum _{i=1}^{N_k} \ln g_{NN, \textit{i}}(\textbf{r},\omega) \right)

\end{equation}
